---
title: "PSTAT126_Homework2"
author: "Yanjie Qi"
date: "2019/8/19"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## 1.
## (a) 
Predictor here is ppgdp and response is fertility.

## (b)
load alr4 package:
```{r}
library(alr4)
data(UN11)
attach(UN11)
```
create two variables from data UN11:
```{r}
fertility=UN11$fertility
ppgdp=UN11$ppgdp  
```
draw scattle plot and the linear model:
```{r}
plot(x=ppgdp,y=fertility)
abline(lm(fertility~ppgdp))
summary(lm(fertility~ppgdp))
```
According to the graph above, The trend is not linear.

## (c)
draw scattle plot and the linear model by using natural log of the variables:
```{r}
plot(x=log(ppgdp),y=log(fertility))
abline(lm(log(fertility)~log(ppgdp)))
summary(lm(log(fertility)~log(ppgdp)))
```
The dependence of response variable (fertility) on predictor (ppgdp) is clearly shown on the log model. Therefore, simple linear regression model is plausible for summary of this graph.

## 2.
load faraway Packages:
```{r}
library(faraway)
names(prostate)
```

## a) 
draw a scatter plot:
```{r}
plot(prostate$lcavol,prostate$lpsa,xlab="lcavol",ylab="lpsa",main="lpsa vs lcavol")
```

We can see that there is an overall positive linear relationship between lspa and lcavol; simple linear regression model seems reasonable.

## b)
access data lpsa and lcavol:
```{r}
y<-prostate$lpsa
x<-prostate$lcavol
```
sample means:
```{r}
xbar<-mean(x)
ybar<-mean(y)
```
sum of sqaures:
```{r}
Sxx<-sum((x-xbar)^2)
Syy<-sum((y-ybar)^2)
Sxy<-sum((x-xbar)*(y-ybar))
```
estimate the value of slope:
```{r}
beta1hat<-Sxy/Sxx
```
estimate the value of intercept:
```{r}
beta0hat<-ybar-beta1hat*xbar
```
Displaying values:
```{r}
sprintf('The Value of xbar is %.4f',xbar)
sprintf('The Value of ybar is %.4f',ybar)
sprintf('The Value of Sxx is %.4f', Sxx)
sprintf('The Value of Syy is %.4f', Syy)
sprintf('The Value of Sxy is %.4f', Sxy)
sprintf('The estimated value of the beta0 is %.4f',beta0hat)
sprintf('The estimated value of the beta1 is %.4f',beta1hat)
sprintf('The estimated regression line is %.4f+%.4fx',beta0hat,beta1hat)
```
calculate the fitted values:
```{r}
yhat<-beta0hat+beta1hat*x
```
draw the fitted line on to the plot from part a):
```{r}
plot(x,y,xlab="lcavol",ylab="lpsa",main="lpsa vs lcavol")
lines(sort(x),yhat[order(x)],col="blue")
```

## c)
get the number of observations:
```{r}
n<-length(x)
```
get the sum of square error:
```{r}
sse<-Syy-beta1hat*Sxy
```
get mean square error, which is the estimate of sigma^2:
```{r}
mse<-sse/(n-2)
```
estimates of stamdard errors:
```{r}
sb1<-sqrt(mse/Sxx)
sb0<-sqrt(mse*sum(x^2)/(n*Sxx))
sprintf('The estimated value of sigma^2 %.4f',mse)
sprintf('The standard error of beta1 is %.4f',sb1)
sprintf('The standard error of beta0 is %.4f',sb0)
```
get the covariance:
```{r}
cov<--mse*xbar/Sxx
sprintf('The estimated covariance between beta0&beta1 is %.4f',cov)
```
t-tests for the null hypotheses beta0 = 0:
```{r}
tb0<-beta0hat/sb0
```
p-value of beta0 = P(T>tb0)+P(T<-tb0):
```{r}
pb0<-pt(abs(tb0),df=n-2,lower.tail=FALSE)+ pt(-abs(tb0),df=n-2,lower.tail=TRUE)
sprintf('The test statistics to test beta0=0 is %.4f, the p-value is %.4f',tb0,pb0)
```
t-tests for the null hypotheses beta1 = 0:
```{r}
tb1<-beta1hat/sb1
```
p-value of beta1 = P(T>tb1)+P(T<-tb1):
```{r}
pb1<-pt(abs(tb1),df=n-2,lower.tail=FALSE)+ pt(-abs(tb1),df=n-2,lower.tail=TRUE)
sprintf('The test statistics to test beta1=0 is %.4f, the p-value is %.4f',tb1,pb1)
```

## 3.
## a)
Loan the data set:
```{r}
data(ftcollinstemp)
attach(ftcollinstemp)
```
draw a scattle plot and the linear model:
```{r}
plot(x=fall,y=winter)
abline(lm(winter~fall))
```
## b)
```{r}
summary(lm(winter~fall))
```
According to Summary, we get p-value is is 0.0428 which is larger than 0.01; therefore, we accept the null hypothesis H0 that the slope is 0 when alpha = 0.01.

## c)
```{r}
PercofVar<-summary(lm(winter~fall))$r.squared
PercofVar
```
We could draw the conclusion from above that the percentage of the variability in winter is explained by fall is around 3.7%.

## 4.
## a)
Loan the data set:
```{r}
data(Heights)
attach(Heights)
```
constructe the linear model:
```{r}
mod <- lm(dheight~mheight)
summary(mod)
```
Given by summary, we first have Intercept 29.91744 and slope 0.54175; Therefore dheight = 29.91744 + 0.54175 * mheight. As shown in the summary above, the standard Error of dheight and mheight is 1.62247 and 0.02596, the coefficient of determination is 0.2408, and the estimate of variance is 2.266^2=5.136167.From the coefficient of determination, 24.08% of variance in dheight is explained by mheight.

## b)
```{r}
confint(mod, 'mheight', level=0.90)
```
a 90% confidence interval for β1 is (0.4990166, 0.5844774)

## c)
```{r}
predict(mod, data.frame(mheight=61), interval="prediction",level=.99)
```
From above, we get a 99% confidence interval for daughters whose mother is 61 inches tall is (57.11531, 68.8127)